---
title: "mlm within-person process challenge and answer key"
author: "Ana DiGiovanni"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load your packages 
```{r}
library(tidyverse)
library(brms)
library(bmlm)
library(bayesplot)
library(tidybayes)
```

# Load your data 

We are going to work again with the PANDA dataset that you have now seen for a few weeks. However, in this challenge we are going to be combining cleaning, reformatting, and modeling from the past 4-weeks, so this should be a challenge! 

Load the data from the SIPPS website: https://columbia-sipps.github.io/workshops/coding_advanced/. As a reminder, these data include a variable for ID and gender, and then assess self-worth (SPPA_SWORTH), communication quality with mothers (MPACS), and generalized anxiety (SCARED_GAD) every six months at 6 different time-points. 

```{r}
d <- read.csv('https://columbia-sipps.github.io/workshop_files/coding-adv-04/mlm-data-cleaning.csv')
```

# Research question 

As opposed to the week when we explored whether/how self-worth changed over time, today we are going to look at the association between communication quality with mothers and self-worth. Specifically, our question is **at timepoints when adolescents reported better communication quality with their mothers (as compared to their own average communication quality across the 6 timepoints) did they concurrently report better/higher self-worth?** Remember, that this is a within-person (as oppossed to a between-person) research question. 

# Subset the data

Keep ID, MPCAS variables and SWORTH variables in the dataset
**hint** use the contains function for MPACS and SWORTH
```{r}
d <- d %>%
  select(ID, contains ("MPACS"), contains ("SWORTH"))
```

# Rename the variables

Make the variable names easier to understand
```{r}
d <- d %>% 
  rename(
    id = 1, 
    communication_1 = 2, 
    communication_2 = 3, 
    communication_3 = 4,
    communication_4 = 5,
    communication_5 = 6,
    communication_6 = 7,
    sworth_1 = 8,   
    sworth_2 = 9,   
    sworth_3 = 10,   
    sworth_4 = 11,   
    sworth_5 = 12,   
    sworth_6 = 13, 
    )
```

# Transform the data to long format

Here, you need to do this separately for self-worth and communication. That is, use the pivot_longer function like we did in the past for time and self-worth, and then create another long dataset for communication. **hint: include include a command for names_prefix to make the time labels the same (and numeric) across the two dataframes you are creating**

You will then need to merge the two dataframes using full_join by both id and time 
```{r}
#self-worth long dataset
d_long_sw <- d %>% 
  tidyr :: pivot_longer(., cols = contains("sworth"), names_to = "time", 
                        names_prefix = "sworth_", values_to = "self_worth")

#communication long dataset
d_long_comm <- d %>% 
  tidyr :: pivot_longer(., cols = contains("comm"), names_to = "time", 
                        names_prefix = "communication_", values_to = "communication")


#merge the two dataframes (using full_join)
#use select to keep the variables we want (id, time, self-worth and communication)
d_long <- full_join(d_long_sw, d_long_comm, by = c("id", "time")) %>%
  select(id, time, self_worth, communication)
```

# Clean your long format data

>- filter out the NA responses for self_worth (use !is.na)
>- filter out the NA responses for communication 

```{r}
d_long <- d_long %>%
  dplyr::filter(!is.na(self_worth)) %>%
  dplyr::filter(!is.na(communication)) 
```

# Rescale time 

Make it so that timepoint 1 is recoded tp be 0, so that time ranges from 0 to 5. 
```{r}
d_long$time <- as.numeric(d_long$time)
d_long$time <- d_long$time - 1
```

# Put communication and self-worth on a 0 to 10 scale 

Self worth was previously on a 1 to 4 scale 
Communication was on a 20 to 100 scale 

There are numerous ways to do this, but one way is to implement this equation: (max new - min new)/(max old - min old) * (variable - max old) + max new 
```{r}
d_long$self_worth <- 10/4 * (d_long$self_worth - 4) + 10
d_long$communication <- 10/100 * (d_long$communication - 100) + 10
```


# Parse within-subject communication from between-subject communication 

**hint, use the isolate function** 
```{r}
d_long <- isolate(d_long, by = "id", value = "communication", which = "both")
```

# Run a bayesian model with random intercept and random slope 

Self-worth should be predicted by time and within-subject communication quality. There should be a random intercept, as well as a random slope for within0subject communication quality. 
```{r, warning=FALSE, results='hide', message=FALSE}
m1 <- brm(self_worth ~ time + communication_cw + (1 + communication_cw | id), data = d_long)

summary(m1)
```

# Write a few sentences to interpret the fixed effects for time and communication
>- Time: self-worth for the typical adolescent decreases slightly over time. Specifically, there is a .13 unit decrease in self-worth (on a 10-point scale) every 6-months. This is statistically significant.
>- Communication: At timepoints when adolescents report higher communication quality with their mothers, as compared to their average communication quality across the study period, there is an associated .40 unit decrease in self-worth (on a 10-point scale). This is also statistically significant. 

# Write a couple sentences to interpret and the random effect of communication quality and quantify heterogeneity 
**hint, when quantifying the heterogeneity interval use the formula: fixed effect +/- (plus or minus) 2 times sd of random effects** 
>- There is significant heterogeneity (variability) in the association between within-person communication quality and self-worth.95% of the slopes fall between the range of -.34 and 1.14 (.40 +/- (2* .37) meaning that for some people, they report WORSE self-worth at times when they report better communication with their mothers, and for other people, they experience a stronger positive association between person-specific increases in communication quality and self-worth than the average participant.  

# Visualize the fixed and random effects 
Plot a spaghetti plot 
```{r}
#pull out the random effects
ranefs <- ranef(m1)  

#make it into a dataframe
ranefs <- data.frame(as.numeric(rownames(ranefs$id)),ranefs$id)

ggplot(d_long, aes(communication_cw, self_worth)) +
  #pulling in person specific intercepts and slopes
      geom_abline(ranefs, 
                  intercept = ranefs$Estimate.Intercept + 8.1807947, 
                  slope = ranefs$Estimate.communication_cw + 0.3985144, 
                  color="black", lwd = .3, alpha = .3) +
  #pulling in fixed effects (average line)
  geom_abline(ranefs, intercept = 8.1807947, slope =  0.3985144, 
              color="red", lwd = 1.5, alpha = 1) +
  xlim(-4, 5) + ylim(0, 10) +
  theme(text=element_text(size=16)) 
```


# Visualize the heterogeneity interval
Make sure to visualize where the fixed effect is as well. 

```{r}
#add the random effects of each person to the fixed effect to get person specific slopes 
ranefs$ind_slopes <- ranefs$Estimate.communication_cw + 0.3985144

ggplot(ranefs, aes(x = ind_slopes)) + geom_histogram(color = "blue", fill = "blue", 
                                                              alpha = .4, bins = 40) +  
  #fixed effect line
  geom_vline(xintercept = 0.3985144, size = 1, color = "black", linetype="dashed") +
  #create the 95% credibility interval (fixed effect +/- 1.96 * sd of random effect)
  geom_vline(xintercept = 0.3985144 - 1.96*.37, size = .5, color = "red") +
  geom_vline(xintercept = 0.3985144 + 1.96*.37, size = .5, color = "red") +
  theme_bw() + theme(text = element_text(size=12)) 
```
