---
title: "SIPPS Coding Workshops (basic track)"
subtitle: "Programming Fundamentals Challenge"
author: "Manasi Jayakumar"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This challenge covers topics covered in the first **two workshops** of the basic coding track. 

This challenge has been adapted from [Columbia University Psychology Department's Scientific Computing Workshops](https://cu-psych-computing.github.io/cu-psych-comp-tutorial/){target="_blank"} and some materials provided by Ana DiGiovanni and Anna Vannucci. 

For each of the following questions, the answer should be expressed as the output of a command or a series of commands.


# Paths

1. Show the file path to R's current working directory on your computer.

```{r}
getwd()
```

2. Write out an **absolute** file path to your Desktop folder on your own computer as a **string**.

```{r}
# On My computer:
"/Users/manas/Desktop"
# Or this, cause tilde abbreviation for home folder behaves as absolute path
"~/Desktop"
```


3. Write out a **relative** file path to your Desktop folder as a **string**, given that you are in R's current working directory.

```{r}
# You may need to use the .. file address to go up one level
"../../../Desktop"
```

# Variables and numeric data

1. Make a variable x that is the sum of 2 plus 2
```{r}
x <- 2 + 2
```


2.	What is 89 cubed?
```{r}
89 ^ 3
```


3.	What is the square root of 8? 
HINT: Look up the following on Google- " R square root function"
```{r}
sqrt(8)
```

4. Create a variable that represents degrees in Fahrenheit; set it to a number corresponding to a temperature of your choosing.
```{r}
# This can be whatever number you please, I chose 72
deg.f <- 72
```

5. Look up the formula for Fahrenheit-to-Celsius conversion, create a new variable corresponding to degrees Celsius, and set its value to the F-to-C conversion of your existing degrees Fahrenheit variable.
```{r}
deg.c <- (deg.f - 32) * (5/9)
```

6. Time to check our math! Create a third variable and set its value to the C-to-F conversion of your newly created degrees Celsius variable.
```{r}
deg.f.from.c <- (9/5) * deg.c + 32
```

7. Output the current value of the first Fahrenheit variable to console.
```{r}
deg.f
```

8. Output the current value of the second Fahrenheit variable to console. Visually verify (just by looking at the numbers for now) that these two values are equal.
```{r}
deg.f.from.c
```


# Vectors and character data
1.	Create two vectors with the following elements: "vector_1" with elements 4, 8, 2, 9 and "vector_2" with elements 2, 4, 1, 3. 
```{r}
vector_1 <- c(4,8,2,9)
vector_2 <- c(2,4,1,3)
```


2. Divide vector 1 by vector 2 and store it in a new variable "y" 
```{r}
y <- vector_1/vector_2

```


3. Find the 3rd element of the variable "y"  (indexing the elements using [ ] )
```{r}
y[3]
```

4. Create a variable "fave_foods" that stores a character vector of length 5. Each element is one of your favorite foods  (e.g. 1st, 2nd, ..., 5th favorite food).
```{r}
fave_foods <- c("chocolate", "icecream", "jalapeno chips", "fries", "watermelon")
```

5. Output the 2nd through 4th elements of "fave_foods" vector to console.
```{r}
fave_foods[2:4]
# Less optimal, but still valid
fave_foods[c(2, 3, 4)]
```

6. Create a variable "is_savory" storing a logical vector of length 5, where the nth element is `TRUE` if your nth favorite food is savory, and `FALSE` if your nth favorite food is sweet. For e.g. if your first favorite food is savory, the first element of "is_savory" will be `TRUE`, otherwise it will be `FALSE`.
```{r}
is_savory <- c(FALSE, FALSE, TRUE, TRUE, FALSE)
```

7. Output all of the savory elements in your fave_foods vector to console. 
HINT 1: you can index a vector with a logical vector, and it will only return the elements that correspond to a "TRUE" value
HINT 2: Try indexing the fave_foods vector using is_savory

```{r}
fave_foods[is_savory]
```
# Relational and logical operators

Run this line of code
```{r}
fav_veggies <- c("carrot", "parsley","lettuce", "kale", "green pepper") #these are my guinea pig's favorite veggies
```

1. Check if the "red pepper" is one of my guinea pig's favorite food 

```{r}
"red pepper" %in% fav_veggies
```

2. Check if "parsley" is one of my guinea pig's favorite food
```{r}
"parsley" %in% fav_veggies
```

3. Is the sum of 57 and 8 greater than the multiplication of 5 and 12? (p.s. don't do this by hand - write the code below)
```{r}
(66+10) > (5*12)
```

Run the following chunk of code 
```{r}
# these are the weights of my guinea pigs in lbs (pounds)
tuco_wt <- 2.6
zuko_wt <- 2.2

# these are the ages of my guinea pigs (in yrs)
tuco_age <- 1.5
zuko_age <- 0.9
```

4. According to my vet, if a guinea is older than a year, and weighs more than 2 lbs, I can give him a particular medication X. Can I give tuco and zuko this medication X? Save the results in a variable as `TRUE` (can give medication) or `FALSE` (cannot give medication) for the two of them. Do this using logical and relational operators (not just looking at the numbers)! 

```{r}
# can you give tuco med?
can_tuco_med <- tuco_age > 1 & tuco_wt > 2
can_tuco_med
```

```{r}
# can you give zuko med? 
can_zuko_med <- zuko_age > 1 & zuko_wt > 2
can_zuko_med
```

5. I go to another vet for a second opinion. According to this second vet, if a guinea pig is older than a year OR weights more than 2 lbs, I can give him the medication X. Can I give my guinea pigs the medication in this case?

```{r}
can_tuco_med2 <- tuco_age > 1 | tuco_wt > 2
can_tuco_med2

can_zuko_med2 <- zuko_age > 1 | zuko_wt > 2
can_zuko_med2
```

# Dataframes

1. Read in shakes.csv into a variable names "shakes". Remember this is the same dataframe we looked at in the workshop with details about plays written by Shakespeare.  
HINT: You will have to load in the package `tidyverse` first.  
```{r}
library(tidyverse)
shakes <- read_csv("shakes.csv")
```

2. Check what types of variables you have in this dataframe
```{r}
shakes
```

3. How many total plays did Shakespeare write?
```{r}
nrow(shakes)
```

4. What genres of plays did he write? How many different genres of plays did he write, according to this data?
HINT: you can use the function unique() to find only unique instances of each genre. 
```{r}
# names of genres of plays
unique(shakes$genre)

# number of genres of plays
length(unique(shakes$genre))
```

5. What is the median number of words in a Shakespeare play? 
HINT: use the function median(), instead of the function mean() we used in the workshop.
```{r}
median(shakes$n.words)
```

