---
title: "SIPPS Coding Workshop: basic track"
subtitle: "07: Interactions in linear models"
date: 'July 20, 2021'
author: Manasi & Camille
output:
  beamer_presentation:
    incremental: FALSE
---

```{r setup, include = FALSE}

library(knitr)
library(psych)
library(dplyr)
library(ggplot2)
library(broom)
library(kableExtra)
library(gridExtra)

knitr::opts_chunk$set(echo = T, warning = F, message = F)
theme_set(theme_bw() + theme(panel.grid = element_blank(), text = element_text(size = 10)))

```

# Linear models: types of effects

> - there are two main classes of effects in linear models: **main effects** and **interactions**

> - so far, the models we've talked about and run in R have all included main effects only: these simply tell you about the effect of *a single variable* on your outcome variable (potentially while controlling for other additional variables)

> - however, often we're interested in **interactions**: cases where the effect of one variable *depends* on the value of another variable

# Linear models: interactions

- let's return to our previous example: where `enjoyment` of restaurant take-out is predicted by two variables: `Yelp rating` and `restaurant type`

- visualizing these variables all together provides some evidence that there might be an interaction
```{r, echo = F, fig.height = 1.9, fig.width = 4, fig.align = "center", tidy = F}

set.seed(123)

# fake data frame
food_df <- data.frame(type = rep(c('pizzeria','taqueria'), each = 10),
                      type_num = rep(c(0.5, 1), each = 10),
                      rating = rep(seq(0.5, 5, by = 0.5), 2))
food_df$enjoyment <- food_df$type_num * food_df$rating + rnorm(nrow(food_df), mean = 0, sd = 0.3)
food_df$enjoyment[food_df$type_num == 1] <- 3 + rnorm(10, mean = 0, sd = 0.5)

food_df_mean <- food_df %>%
  group_by(type) %>%
  summarise(mean_enjoyment = mean(enjoyment))

# plot
p2 <- ggplot(data = food_df, aes(x = rating, y = enjoyment, color = type)) +
  geom_point()
p2

```

# Linear models: interactions

> - this previous example shows a case of a **2-way continuous-by-categorical interaction**

> - it allows us to model the fact that there may be *different slopes of our continuous variable (Yelp rating)* for *each level of our categorical variable (restaurant type)*

> - unsurprisingly, you can also have **continuous-by-continuous** interactions and **categorical-by-categorical** interactions

> - we'll go through all three of these types of models today
